---
title: "Quadvariate Viz"
execute: 
  echo: true 
  warning: false
  error: false 
  
format: 
  html:
   code-fold: true 
---

Use this file to generate a professional looking **quadvariate** visualization.  The visualization will not perfect the first time but you are expected to improve on it throughout the semester especially after covering advanced topics such as effective viz.

```{r}
# Load Libraries
library(tidyverse)
library(fivethirtyeight)
library(ggplot2)
```


```{r}
### QUADVARIATE VIZUALIZATION ###

# Create `new_bechdel` Dataset
new_bechdel <- bechdel |>
  mutate(clean_test = factor(clean_test, c("nowomen", "notalk", "men", "dubious", "ok"))) |>
  mutate(half_decades = cut(year, breaks = seq(1969, 2014, by = 5)))
```

```{r}
### CREATE PLOT ###  

ggplot(new_bechdel, aes(x = budget, y = domgross, color = binary)) +
  geom_point(alpha = 0.7, size = 0.5) +
  geom_smooth(method = "lm") + 
  facet_wrap(~ clean_test,  labeller = labeller(clean_test = c("nowomen" = "No Women",
                                                "notalk" = "No Talk",
                                                "men" = "Men Only",
                                                "dubious" = "Dubious",
                                                "ok" = "Passes")),
             ncol = 3, # Specify the number of columns for the facets
             scales = "free_x") + # Allow x-axis scales to vary if needed +
   labs(
    title = "Bechdel Test Results by Reason for Failure",
    x = "Budget (in USD)",
    y = "Domestic Gross (in USD)",
    caption = "Source: Walt Hickey, 'The Dollar-And-Cents Case Against Hollywood’s Exclusion of Women,' FiveThirtyEight, 2014",
    color = "Bechdel Test Result") +
  theme_minimal() + 
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 16),
    axis.title.x = element_text(face = "bold", size = 12),
    axis.title.y = element_text(face = "bold", size = 12),
    legend.title = element_text(face = "bold", size = 12),
    strip.text = element_text(face = "bold", size = 11),
    plot.caption = element_text(size = 8, hjust = 0.5), 
    panel.spacing.x = unit(1, "cm"),
    panel.spacing.y = unit(1, "cm"), 
    axis.text.x = element_text(angle = 45, hjust = 1, size = 10), 
    axis.text.y = element_text(size = 10),
    plot.margin = margin(1, 1, 1, 1, "cm"))
```

# Reflection 

This quadvariate visualization uses a faceted scatter plot to demonstrate the relationship between a movie's Budget (x-axis) and its Domestic Gross (y-axis), segmented by the Bechdel Test Result (color: red for FAIL, blue-green for PASS) and further faceted by the 'Reason for Failure' on the Bechdel Test (No Women, No Talk, Men Only, Dubious) and a category for movies that Pass.

The most important takeaway is that there doesn’t appear to be a strong or direct financial reward or penalty tied to whether a film passes or fails the Bechdel Test in this dataset. Both high-grossing and low-grossing films can be found among those that pass the test as well as among the various failure categories. While there is a general trend showing that films with larger budgets tend to earn more, this pattern seems consistent regardless of Bechdel Test results. 
