---
title: "Trivariate Viz"
execute: 
  echo: true 
  warning: false
  error: false 
  
format: 
  html:
   code-fold: true 
---

Use this file to generate a professional looking **trivariate** visualization.  The visualization will not perfect the first time but you are expected to improve on it throughout the semester especially after covering advanced topics such as effective viz.

```{r}
### TRIVARIATE VIZUALIZATION ###

# Load Libraries 
library(ggplot2)
library(knitr)
library(kableExtra)

# Import Education Data
education <- read.csv("https://mac-stat.github.io/data/sat.csv")

# Create Table Previewing The Dataset 
kable(head(education), caption = "Education Data Sample") |> 
  kable_styling(bootstrap_options = c("striped", "hover"))
```

```{r}
# Download Library 
library(ggplot2)

# Create Density Plot
ggplot(education, aes(x = sat, fill = fracCat)) + 
  geom_density(alpha = 0.6) +                                                # Adjust Transparency 
  labs(
    title = "SAT Score Distributions by Student Participation Rate",
    x = "SAT Total Score",
    y = "Density",
    caption = "Source: College Board (compiled by Danny Kaplan, 2021)",
    fill = "SAT Participation Rate (%)") +
  theme_minimal() +                                                           # Cleaner theme
  theme(
    plot.title = element_text(face = "bold", size = 15, hjust = 0.5),         # Make Title Bold and Centered
    plot.caption = element_text(size = 10, hjust = 0.5),                      # Adjust size of caption
    panel.grid.major = element_line(color = "gray80"),                        # Adjust grid lines
    legend.position = "right")                                                # Position the Legend on the right
```

# Reflection

This trivariate overlapping density plot illustrates the distribution of SAT total scores (x-axis) by density (y-axis), segmented by three color-coded ranges of student participation rates (legend). It allows for a clear comparison of score distributions across states with varying levels of SAT participation.

The most striking takeaway is the clear inverse relationship between the percentage of students taking the SAT and the overall distribution of SAT scores in a state. In states with low participation, the average scores tend to be inflated. Conversely, in states with much higher participation, the average scores tend to be lower.

```{r}
# Create A Scatter Plot with Linear Regression Lines
library(ggplot2)
ggplot(education, aes(x = expend, y = sat, color = fracCat)) +
  geom_point(alpha = 0.7, size = 3) +                                         #Adjust Transparency and Point Size
  geom_smooth(method = "lm", se = TRUE, linewidth = 1) +                      # Keep regression lines with confidence intervals
  labs(
    title = "Expenditure Per Student and SAT Scores by Participation Rate",
    x = "Expenditure Per Student",
    y = "SAT Total Score",
    caption = "Source: College Board (compiled by Danny Kaplan, 2021)",
    color = "SAT Participation Rate") +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold", size = 12, hjust = 0.5),         # Make Title Bold and Centered
    plot.caption = element_text(size = 10, hjust = 0.5),                      # Adjust size of caption
    panel.grid.major = element_line(color = "gray80"),                        # Adjust grid lines
    legend.position = "right")                                                # Position the Legend on the right
```

# Reflection 

This trivariate scatter plot displays the relationship between two numerical variables, Expenditure Per Student (on the x-axis) and SAT Total Score (on the y-axis). The third variable, SAT Participation Rate, is introduced through the color of the data points and the separate linear regression lines. 

The plot shows that the proportion of students taking the SAT seems to be a stronger predictor of average SAT scores than expenditure per student. States with lower SAT participation rates tend to have higher average scores, regardless of their spending per student. Conversely, states with high participation rates tend to have lower average scores.


