---
title: "Solo Project"
execute: 
  echo: true 
  warning: false
  error: false 
  
format: 
  html:
   code-fold: true 
---

For my solo project, I wanted to visualize the shifts in Democratic voteshare across Minnesota’s eight congressional districts between the 2022 midterm elections and the 2024 general election. This project involved collecting certified election results from both cycles, calculating the change in Democratic voteshare, and displaying those shifts through a comparative visual format.

```{r}
# Load Necessary Libraries
library(tigris)
library(leaflet)
library(sf)
library(dplyr)
library(tidyr)
library(tidyverse)
library(readr)
library(stringr)
library(knitr)
library(kableExtra)
```

```{r}
# Load United States Congressional Districts
cd118 <- congressional_districts(cb = TRUE, resolution = "20m", year = 2022, progress_bar = FALSE)

# Create Map
leaflet(cd118) |>
  addTiles() |>
  addPolygons()
```

```{r}
# Define Minnesota Congressional Districts
mn_cd <- congressional_districts(state = "MN", cb = TRUE, resolution = "20m", year = 2022)

# Create Map
leaflet(mn_cd) |>
  addTiles() |>
  addPolygons(
    label = ~paste("District", CD118FP),)
```

```{r}
#Import 2022 Minnesota Congressional Results 
data_2022 <- read_delim("../data/Election_Results_2022", col_names = FALSE)

# Clean and filter data
Minnesota_Congressional_Results_2022 <- data_2022 |>
  select(X6, X11, X15) |>
  rename(District = X6,
         Party = X11,
         Voteshare = X15) |>
  filter(Party %in% c("DFL", "R"))  # Filter for DFL and R parties

Minnesota_Congressional_Results_2022

# Show Table
kable(head(Minnesota_Congressional_Results_2022), 
      caption = "Minnesota Congressional Election Results 2022") |>
  kable_styling(bootstrap_options = c("striped", "hover"))
```

```{r}
#Import 2024 Minnesota Congressional Results 
data_2024 <- read_delim("../data/Election_Results_2024", col_names = FALSE)

Minnesota_Congressional_Results_2024 <- data_2024 |>
  select(X6, X11, X15) |>
  rename(District = X6,
         Party = X11,
         Voteshare = X15)  |>
  filter(Party %in% c("DFL", "R"))  # Keep only DFL and R parties 

# Show Table
kable(head(Minnesota_Congressional_Results_2024), 
      caption = "Minnesota Congressional Election Results 2024") |>
  kable_styling(bootstrap_options = c("striped", "hover"))
```

```{r}
#Combine 2022 and 2024 Data
Minnesota_Congressional_Results_2022 <- Minnesota_Congressional_Results_2022 |>
  mutate(Year = 2022)

Minnesota_Congressional_Results_2024 <- Minnesota_Congressional_Results_2024 |>
  mutate(Year = 2024)

Combined_Minnesota_Congressional_Results <- bind_rows(
  Minnesota_Congressional_Results_2022,
  Minnesota_Congressional_Results_2024)

# Fix District Variable (To Match Geographic Data)
Combined_Minnesota_Congressional_Results$District <- sprintf("%02d", Combined_Minnesota_Congressional_Results$District)

# Show Table
kable(head(Combined_Minnesota_Congressional_Results), 
      caption = "Combined Congressional Results") |>
  kable_styling(bootstrap_options = c("striped", "hover"))
```

```{r}
# Create the 'year_party' column for pivoting
Combined_Minnesota_Congressional_Results <- Combined_Minnesota_Congressional_Results |>
  mutate(year_party = paste0(tolower(Party), "_", Year))

# Pivot The Data
voteshare_data <- Combined_Minnesota_Congressional_Results |>
  pivot_wider(id_cols = District,
              names_from = year_party,
              values_from = `Voteshare`)

# Show Table
kable(head(voteshare_data, 8), 
      caption = "Voteshare Data") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

```{r}
#Merge Spatial and Numeric Data
mn_cd_with_results <- left_join(mn_cd, voteshare_data, by = c("CD118FP" = "District"))
```

```{r}
# Create Color Palette for 2022 
pal_2022 <- colorNumeric(palette = c("red", "blue"), domain = mn_cd_with_results$democratic_2022)

# Map for 2022 Election Results
leaflet(mn_cd_with_results) |> 
  addTiles() |> 
  addPolygons(
    fillColor = ~pal_2022(dfl_2022),  # Function needs to be called on the variable
    fillOpacity = 1,
    color = "black",
    weight = 1) |> 
  addLegend(
    pal = pal_2022,
    values = ~dfl_2022,
    title = "Democratic Voteshare (2022)",
    position = "bottomright")
```

```{r}
# Create Color Palette for 2024
pal_2024 <- colorNumeric(palette = c("red", "blue"), domain = mn_cd_with_results$democratic_2022)

# Map for 2024 Election Results
leaflet(mn_cd_with_results) |> 
  addTiles() |> 
  addPolygons(
    fillColor = ~pal_2024(dfl_2024),
    fillOpacity = 1,
    color = "black",
    weight = 1) |> 
  addLegend(
    pal = pal_2024,
    values = ~dfl_2024,
    title = "Democratic Voteshare (2024)",
    position = "bottomright")
```

```{r}
# Calculate the Change in Voteshare (Between 2022 and 2024)
mn_cd_with_results_1 <- mn_cd_with_results |>
  mutate(
    democratic_change = dfl_2024 - dfl_2022,
    republican_change = r_2024 - r_2022)
```

```{r}
# Map for Democratic Voteshare change
plot_dem_change <- ggplot(mn_cd_with_results_1) +
  geom_sf(aes(fill = democratic_change), color = "black") +
  scale_fill_gradient2(
    low = "red",
    mid = "white",
    high = "blue",
    midpoint = 0,
    name = "Dem. Change\n(2022 - 2024)") +
  labs(
    title = "Change in Democratic Voteshare (2022 to 2024)",
    caption = "Source: Minnesota Secretary of States Office (2023 and 2025)") +
  theme_minimal() + 
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank())

plot_dem_change
```

# Reflection: 

The map offers several insightful observations about the shifts in Democratic voteshare between the 2022 midterm elections and the 2024 elections. Notably, Minnesota's 2nd Congressional District experienced a significant shift in favor of Democratic candidate Angie Craig. This district has long been considered a swing district, but in 2024, Craig secured a commanding victory, greatly surpassing her 2022 performance and outperforming other Democrats in the state. Additionally, while Minnesota's 7th District remains a heavily Republican area, it did shift slightly to the left. Most other districts showed moderate shifts to the right, consistent with national trends. Interestingly, Minnesota's 5th District, the most Democratic in the state, showed no change between the two election cycles.

These shifts provide valuable insights into where Democratic candidates gained or lost support in 2024 and open the door to several important research questions. What campaign strategies, narratives, or tactics did certain candidates employ that may have contributed to their increased support? What role did money play in the outcome? These are all compelling questions that warrant further exploration by political scientists and data scientists alike.

