---
title: "Spatial Viz"
execute: 
  echo: true 
  warning: false
  error: false 
  code-fold: true 
---

Use this file for practice with the **spatial viz** in-class activity.  Refer to the class website for details.

```{r}
### Exercise 1: A Leaflet With Markers / Points ###

fave_places <- read.csv("https://hash-mac.github.io/stat112site-s25/data/our_fave_places.csv")

# Check it out
head(fave_places)
```

```{r}
### Part A ###

# Load Leaflet
library(leaflet)

# Just a plotting frame
leaflet(data = fave_places)
```

```{r}
# Now what do we have?
leaflet(data = fave_places) |> 
  addTiles() |> 
  addMarkers(lng = ~longitude, lat = ~latitude)
```


```{r}
### Exercise 2: Details ###

# Load package needed to change color
library(gplots)

# We can add colored circles instead of markers at each location
leaflet(data = fave_places) |> 
  addTiles() |> 
  addCircles(color = col2hex("red"))
```

```{r}
leaflet(data = fave_places) |>
  addProviderTiles("USGS") |> # Change the background 
  addCircles(weight = 10, opacity = 1, color = col2hex("yellow")) |> #Mark locations with yellow dots
  addPolylines(
    lng = ~longitude,
    lat = ~latitude,
    color = col2hex("green")) # Connect the dots with green lines
```

```{r}
### Exercise 3 ###

#Load Libraries
library(dplyr)
library(leaflet)
library(ggplot2)

# Import starbucks location data
starbucks <- read.csv("https://mac-stat.github.io/data/starbucks.csv")

# Filter for Starbucks in Minnesota
starbucks_mn <- starbucks |>   
  filter(Country == "US", State.Province == "MN")

#Create Leaflet
leaflet(data = starbucks_mn) |> 
  addTiles() |> 
  addMarkers()
```

```{r}
### Exercise 4 ###

#Part A: First, we can grab country-level boundaries from the rnaturalearth package.

# Load the rnaturalearth package
library(rnaturalearth)

# Retrieve world country boundaries in "sf" (simple features) format
world_boundaries <- ne_countries(returnclass = "sf")
```

```{r}
#Part B
# This code produces a world map showing country boundaries.
# What geom are we using for the point map? We are using geom_sf() to create the geometric shapes 
ggplot(world_boundaries) + 
  geom_sf()
```

```{r}
# Load package needed to change map theme
library(mosaic)

# Add a point for each Starbucks
# NOTE: The Starbucks info is in our starbucks data, not world_boundaries
# How does this change how we use geom_point?!
ggplot(world_boundaries) + 
  geom_sf() + 
  geom_point(
    data = starbucks,
    aes(x = Longitude, y = Latitude),
    alpha = 0.3, size = 0.2, color = "darkgreen") +
  theme_map()
```

Part C: The map visually highlights that Starbucks has a strong presence in North America, Europe, and East Asia, a moderate presence in some other regions, and a very limited footprint in Africa.

```{r}
### Exercise 5: Zooming in on some countries ###

#Part A
starbucks_cma <- starbucks |> 
  filter(Country %in% c('CA', 'MX', 'US'))

#A Background Map of State
cma_boundaries <- ne_states(
  country = c("canada", "mexico", "united states of america"),
  returnclass = "sf")
```

```{r}
#Part B
# Just the boundaries
ggplot(cma_boundaries) + 
  geom_sf()
```

```{r}
# Add the points
# And zoom in
ggplot(cma_boundaries) + 
  geom_sf() + 
  geom_point(
    data = starbucks_cma,
    aes(x = Longitude, y = Latitude),
    alpha = 0.3,
    size = 0.2,
    color = "darkgreen"
  ) +
  coord_sf(xlim = c(-179.14, -50)) +
  theme_map()
```

```{r}
# Exercise 6: Create a state and county-level map

#Part A
starbucks_midwest <- starbucks |> 
  filter(State.Province %in% c("MN", "ND", "SD", "WI"))

# Load packages
library(sf)
library(maps)

# Get the boundaries
midwest_boundaries <- st_as_sf(
  maps::map("county",
            region = c("minnesota", "wisconsin", "north dakota", "south dakota"), 
            fill = TRUE, plot = FALSE))

# Check it out
head(midwest_boundaries)
```

```{r}
# Part B
ggplot(midwest_boundaries) +
  geom_sf() +
  geom_point(
    data = starbucks_midwest,
    aes(x = Longitude, y = Latitude),
    alpha = 0.7,
    size = 0.2,
    color = 'darkgreen') +
  theme_map()
```


```{r}
### Exercise 7: Contour Maps ###

# Point map (we made this earlier)
ggplot(cma_boundaries) + 
  geom_sf() + 
  geom_point(
    data = starbucks_cma,
    aes(x = Longitude, y = Latitude),
    alpha = 0.3,
    size = 0.2,
    color = "darkgreen"
  ) +
  coord_sf(xlim = c(-179.14, -50), ylim = c(14.54, 83.11)) +
  theme_map()
```
```{r}
# What changed in the plot?
# What changed in our code?!
ggplot(cma_boundaries) + 
  geom_sf() + 
  geom_density_2d(
    data = starbucks_cma,
    aes(x = Longitude, y = Latitude),
    size = 0.2,
    color = "darkgreen"
  ) +
  coord_sf(xlim = c(-179.14, -50), ylim = c(14.54, 83.11)) +
  theme_map()
```
```{r}
### PART 3: Choropleth Maps ### 
elections_by_state <-  read.csv("https://mac-stat.github.io/data/election_2020_by_state.csv")
elections_by_counties <- read.csv("https://mac-stat.github.io/data/election_2020_county.csv")
```

```{r}
#CODE
elections_by_state <- elections_by_state |> 
  filter(state_abbr != "DC") |> 
  select(state_name, state_abbr, repub_pct_20) |> 
  mutate(repub_20_categories = 
           cut(repub_pct_20, 
               breaks = seq(30, 70, by = 5), 
               labels = c("30-34", "35-39", "40-44", "45-49",
                          "50-54", "55-59", "60-64", "65-70"), 
               include.lowest = TRUE))

elections_by_counties <- elections_by_counties |> 
  select(state_name, state_abbr, county_name, county_fips,
          repub_pct_20, median_age, median_rent) |> 
  mutate(repub_20_categories = 
           cut(repub_pct_20, 
               breaks = seq(0, 100, by = 10),
               labels = c("0-9", "10-19", "20-29", "30-39", "40-49",
                          "50-59", "60-69", "70-79", "80-89", "90-100"),
               include.lowest = TRUE))
```

Exercise 8: State-level choropleth maps

```{r}
#Part A

# Get the latitude and longitude coordinates of state boundaries
states_map <- map_data("state")

# Check it out
head(states_map)
```

```{r}
# Part B
# Note where the dataset, elections_by_state, is used
# Note where the background map, states_map, is used
ggplot(elections_by_state, aes(map_id = state_name, fill = repub_pct_20)) +
  geom_map(map = states_map) +
  expand_limits(x = states_map$long, y = states_map$lat) +
  theme_map() 
```

```{r}
# Make it nicer!
ggplot(elections_by_state, aes(map_id = state_name, fill = repub_pct_20)) +
  geom_map(map = states_map) +
  expand_limits(x = states_map$long, y = states_map$lat) +
  theme_map() + 
  scale_fill_gradientn(name = "% Republican", colors = c("blue", "purple", "red"), values = scales::rescale(seq(0, 100, by = 5)))
```

```{r}
ggplot(elections_by_state, aes(map_id = state_name, fill = repub_20_categories)) +
  geom_map(map = states_map) +
  expand_limits(x = states_map$long, y = states_map$lat) +
  theme_map()
```

```{r}
# Load package needed for refining color palette
library(RColorBrewer)

# Now fix the colors
ggplot(elections_by_state, aes(map_id = state_name, fill = repub_20_categories)) +
  geom_map(map = states_map) +
  expand_limits(x = states_map$long, y = states_map$lat) +
  theme_map() + 
  scale_fill_manual(values = rev(brewer.pal(8, "RdBu")), name = "% Republican")
```

```{r}
#Part C
# Get only the starbucks data from the US
starbucks_us <- starbucks |> 
  filter(Country == "US")

# Map it
ggplot(elections_by_state, aes(map_id = state_name, fill = repub_20_categories)) +
  geom_map(map = states_map) +
  geom_point(
    data = starbucks_us,
    aes(x = Longitude, y = Latitude),
    size = 0.05,
    alpha = 0.2,
    inherit.aes = FALSE
  ) +
  expand_limits(x = states_map$long, y = states_map$lat) +
  theme_map() + 
  scale_fill_manual(values = rev(brewer.pal(8, "RdBu")), name = "% Republican")
```

Exercise 9: County-level choropleth maps

```{r}
#Part A
# Get the latitude and longitude coordinates of county boundaries
library(socviz)
data(county_map) 

# Check it out
head(county_map)
```

```{r}
# Add 0's at the beginning of any fips_code that's fewer than 5 numbers long
# Don't worry about the syntax
elections_by_counties <- elections_by_counties |> 
  mutate(county_fips = as.character(county_fips)) |> 
  mutate(county_fips = 
           ifelse(nchar(county_fips) == 4, paste0("0", county_fips), county_fips))
```

```{r}
#Part B
ggplot(elections_by_counties, aes(map_id = county_fips, fill = repub_20_categories)) +
  geom_map(map = county_map) +
  scale_fill_manual(values = rev(brewer.pal(10, "RdBu")), name = "% Republican") +
  expand_limits(x = county_map$long, y = county_map$lat) +
  theme_map() +
  theme(legend.position = "right") + 
  coord_equal()
```

### Spacial Vizualization: Exercise #10 ###
Construct county-level maps of median_rent and median_age.


```{r}
# Download Datasets
elections_by_state <-  read.csv("https://mac-stat.github.io/data/election_2020_by_state.csv")
elections_by_counties <- read.csv("https://mac-stat.github.io/data/election_2020_county.csv")
```

```{r}
# Wrangle the Data
elections_by_state <- elections_by_state |> 
  filter(state_abbr != "DC") |> 
  select(state_name, state_abbr, repub_pct_20) |> 
  mutate(repub_20_categories = 
           cut(repub_pct_20, 
               breaks = seq(30, 70, by = 5), 
               labels = c("30-34", "35-39", "40-44", "45-49",
                          "50-54", "55-59", "60-64", "65-70"), 
               include.lowest = TRUE))

elections_by_counties <- elections_by_counties |> 
  select(state_name, state_abbr, county_name, county_fips,
          repub_pct_20, median_age, median_rent) |> 
  mutate(repub_20_categories = 
           cut(repub_pct_20, 
               breaks = seq(0, 100, by = 10),
               labels = c("0-9", "10-19", "20-29", "30-39", "40-49",
                          "50-59", "60-69", "70-79", "80-89", "90-100"),
               include.lowest = TRUE))
```

```{r}
# Get Latitude and Longitude Coordinates of States
states_map <- map_data("state")

# Check It Out
head(states_map)
```
```{r}
# Get the latitude and longitude coordinates of county boundaries
library(socviz)
data(county_map) 

# Check it out
head(county_map)
```

```{r}
# Create A County-Level Map of median_rent 
ggplot(elections_by_counties, aes(map_id = county_fips, fill = median_rent)) +
  geom_map(map = county_map) +
  expand_limits(x = county_map$long, y = county_map$lat) +
  theme_map() +
  theme(legend.position = "right") + 
  coord_equal() + 
  scale_fill_gradientn(name = "median rent", colors = c("white", "lightgreen", "darkgreen"))
```

