---
title: "Data Import"
execute: 
  echo: true 
  warning: false
  error: false 
  code-fold: true 
---

Exercise 2: Import Data to RStudio

```{r}
#Load Libraries
library(readr)
library(dplyr)
library(tidyverse)

#Load Data (Relative File Path)
imdb_messy <- read_csv("../data/imdb_5000_messy.csv")
```
Exercise #3: Check Data

```{r}
#Part B
imdb_messy |>
  mutate(across(where(is.character), as.factor)) |> 
  summary()
```

Exercise 4: Clean Data: Factor Variables 1

```{r}
#Count Color Variable
imdb_messy |> 
  count(color)
```

Exercise 5: Clean Data: Factor Variables 2

Part A: 

It's important to use RCode because we can reproduce it. If we clean the data manually in a spreadsheet, there’s no record of what we changed. By using R code and saving it in the qmd -- we can show our work so that anyone else could understand what we did and how we got there. 

```{r}
#Part B
imdb_temp <- imdb_messy |> 
  mutate(color = fct_recode(color,
                            "Color" = "COLOR",
                            "Color" = "color",
                            "Black_White" = "B&W",
                            "Black_White" = "Black and White"))

imdb_temp |> 
  count(color)
```


```{r}
#Part C
imdb_temp <- imdb_messy |> 
  mutate(color = str_replace(color, "COLOR", "Color"),
         color = str_replace(color, "color", "Color"),
         color = str_replace(color, "B&W", "Black_White"),
         color = str_replace(color, "Black and White", "Black_White"))

imdb_temp |> 
  count(color)
```

Exercise 6: Clean Data: Missing Data 1

```{r}
#Part A
nrow(imdb_messy)
```

```{r}
colSums(is.na(imdb_messy))
```

```{r}
#Part C
imdb_messy |> 
  filter(is.na(actor_1_facebook_likes))
```

Exercise 7: Clean: Missing Data 2

```{r}
#Part A
imdb_messy |> 
  summarize(mean(duration, na.rm = TRUE))
```

```{r}
#Part B
imdb_temp <- drop_na(imdb_messy)
nrow(imdb_temp)
```

```{r}
colSums(is.na(imdb_temp))
```

```{r}
#Part C
ggplot(imdb_messy, aes(x = duration, fill = color)) +
  geom_density()
```

